An Efficient Feedback Calibration Algorithm for Direct Imaging Radio Telescopes

### 2016-06-08 cal_demo/ms.tex
Updated notation in section 5

For each simulation run we find gain solutions using four methods:
\begin{enumerate}[i.]
\item EPICal, using a full sky model
\item EPICal, using a single point source model
\item Visibility-based, using a full sky model
\item Visibility-based, using a single point source model.
\end{enumerate}
In the cases where we use the full sky model, the model visibilities, $\V'_{ab}$ used in the 
calibration loop are created using all point sources in the simulated sky. For the single point 
source model, we only allow the calibration loop to be privy to the bright calibrator source. 
While we vary the model used for calibration, the simulated ``true sky'' remains fixed with all 50 
point sources.

For the EPICal calibration loops, we initialize our gain estimates with the true values 
(${g^{(0)}_a=g^T_a=1}$), and allow the estimate to be corrupted by the noise through ten 
iterations of the calibration loop. We adopt a damping factor $\damp=0.35$. Because EPICal 
updates the gain estimate at each calibration loop, \tcal, but retains a memory of previous 
iterations through the damping factor, the total integration time is not straightforward. We define 
an effective integration time by considering the relative weights of each previous $\Cna$ 
contributing to the current $g^{(n)}_a$ estimate, each with integration time \tcal. In the limit $n
\rightarrow \infty$, the geometric series converges to an effective integration time of
\begin{equation}
\teff = \tcal \times \frac{1+\damp}{1-\damp}.
\end{equation}
With the damping factor we adopted here, $\teff \approx 2.1 \times \tcal$.

We simultaneously form simulated visibilities by correlating all pairs of antenna E-field 
measurements (including the receiver noise). The E-fields are correlated for a duration equal to 
the effective integration time of the EPICal loop for comparison. We then find the visibility-
based gain estimates by minimizing
\begin{equation}\label{eq:vis_cal}
\chi^2 = \sum_a\sum_{b\ne a} \left|\V_{ab}-g_a g_b^* \V'_{ab}\right|^2,
\end{equation}
using both versions of the model visibilities described above.

 For each version of calibration, we observe the error in the gain estimates averaging over both 
 antennas and frequency channels.
\begin{equation}\label{eq:gain_error}
\sigma_g = \left[\frac{1}{N_f \Nant} \sum_f \sum_a \frac{\left|g'_a(f)-g^T_a(f)\right|^2}{\left|g^T_a(f)\right|^2}\right]^{1/2}
\end{equation}

>>>>

For each simulation run we find gain solutions using four methods:
\begin{enumerate}[i.]
\item EPICal, using a full sky model
\item EPICal, using a single point source model
\item Visibility-based, using a full sky model
\item Visibility-based, using a single point source model.
\end{enumerate}
In the cases where we use the full sky model, the model visibilities, $V^M_{ab}$ used in the 
calibration loop are created using all point sources in the simulated sky. For the single point 
source model, we only allow the calibration loop to be privy to the bright calibrator source. 
While we vary the model used for calibration, the simulated ``true sky'' remains fixed with all 50 
point sources.

For the EPICal calibration loops, we initialize our gain estimates with the true values 
(${g^{(0)}_a=g_a=1}$), and allow the estimate to be corrupted by the noise through ten 
iterations of the calibration loop. We adopt a damping factor $\damp=0.35$. Because EPICal 
updates the gain estimate at each calibration loop, \tcal, but retains a memory of previous 
iterations through the damping factor, the total integration time is not straightforward. We define 
an effective integration time by considering the relative weights of each previous $\Cna$ 
contributing to the current $g^{(n)}_a$ estimate, each with integration time \tcal. In the limit $n
\rightarrow \infty$, the geometric series converges to an effective integration time of
\begin{equation}
\teff = \tcal \times \frac{1+\damp}{1-\damp}.
\end{equation}
With the damping factor we adopted here, $\teff \approx 2.1 \times \tcal$.

We simultaneously form simulated visibilities by correlating all pairs of antenna E-field 
measurements (including the receiver noise). The E-fields are correlated for a duration equal to 
the effective integration time of the EPICal loop for comparison. We then find the visibility-
based gain estimates by minimizing
\begin{equation}\label{eq:vis_cal}
\chi^2 = \sum_a\sum_{b\ne a} \left|V_{ab}-g_a g_b^* V^M_{ab}\right|^2,
\end{equation}
using both versions of the model visibilities described above.

 For each version of calibration, we observe the error in the gain estimates averaging over both 
 antennas and frequency channels.
\begin{equation}\label{eq:gain_error}
\sigma_g = \left[\frac{1}{N_f \Nant} \sum_f \sum_a \frac{\left|g^{(n)}_a(f)-g_a(f)\right|^2}{\left|g_a(f)\right|^2}\right]^{1/2}
\end{equation}
